---
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

```

# h3r

**h3r** is an interface to [{h3lib}](https://github.com/SymbolixAU/h3lib), which is itself a wrapper around Uber's [H3](https://h3geo.org/) library. See their [getting started](https://h3geo.org/docs/) guide for all the details.

The wrappers are all vectorised, meaning each input can take a vector, and / or return a vector. 

e.g:

```{r}
h3r::latLngToCell(
  lat = c(-37.820197, -37.818476)
  , lng = c(144.983324, 144.967354)
  , resolution = c(1L, 14L)
  )

```


Most of the [H3 API](https://h3geo.org/docs/api/indexing) is included in this package as R functions. The only exceptions are

- stringToH3
- h3ToString
- gridDisksUnsafe
- cellToChildrenSize
- uncompactCellsSize
- maxPolygonToCellsSize
- cellsToLinkedMultipolygon
- destroyLInkedMultiPolygon

However, these should all be accessible through the C API

## Design Choices

- There is no `h3` class, or nice printing, or any fancy sugar-coating of what's returned from the functions. I've kept the outputs as raw / primitive as possible
- All H3Indexes are returned as the String representation. If you want the `H3Index` / `uint64_t` type you need to use the C / C++ functions directly
- We're using the same code conventions as per the H3 API 
  - CamelCase function names (e.g. `latLngToCell`)
  - lat/lng, rather than lat/lon or lon/lat
- Functions which return a single coordinate pair for each input return a `data.frame`


```{r}
h3r::cellToLatLng(cell = c("8cbe63562a54bff","8cbe635631103ff"))
```

- Functions which return a multiple coordinates for each input return a list of `data.frames`


```{r}
h3r::cellToBoundary(cell = c("8cbe63562a54bff","8cbe635631103ff"))
```


## API

To use the source C / C++ code in your own package you should only need to include `inst/include/h3rapi.h`



## C API

In `/inst/capi` you'll find a demo package {h3rc}. This shows how to include / call the C code from {h3r} into another package.

The main components you need to address are

1. DESCRIPTION 
2. `src/init.c`
3. `src/myCode.c` - i.e., your own C code
4. `R/myRCode.R`  - i.e., your own R code
5. Register the dynamic library

---


### DESCRIPTION

Depend & Link to {h3r}

```
Depends:
  h3r
LinkingTo:
  h3r
```

### src/init.c

Define the functions you want to import from {h3r}

```c
SEXP (*h3rLatLngToCell)(SEXP,SEXP,SEXP);

void R_init_h3rc(DllInfo *info)

{
  R_registerRoutines(info, NULL, callMethods, NULL, NULL);
  R_useDynamicSymbols(info, FALSE);

  /* Imports from h3r */
  h3rLatLngToCell = (SEXP(*)(SEXP,SEXP,SEXP)) R_GetCCallable("h3r", "h3rLatLngToCell");
}
```

### src/h3rc.c

Include the "h3rapi.h" header, then you can call the functions you've registered in `src/init.c`

```c
#include "h3rapi.h"

SEXP h3rcLatLngToCell(SEXP lat, SEXP lng, SEXP res) {
  return h3rLatLngToCell(lat, lng, res);
}

```

### R/h3rr.R

Call the function you've defined in `h3rc.c` from within an R function

```r
#' @export
ll2Cell <- function(lat, lon, res) {
  .Call(h3rcLatLngToCell, lat, lon, res)
}

```


### R/<somewhere.R>

Register your dynamic routines. If using Roxygen to build and document your pacakge you can specify

```r
#' @useDynLib h3rc, .registration = TRUE
NULL
```

and it will be built and added to your NAMESPACE automatically



## C++ API

In `/inst/cppapi` you've find a demo package {h3rcpp}. This shows how to include / call the C++ code from {h3r} into another package.

The main components you need to address are

1. DESCRIPTION 
3. `src/myCode.cpp` - i.e., your own C++ code
4. `R/myRCode.R`  - i.e., your own R code
5. Register the dynamic library

The example package shows how to do this, and it's very similar to the `C` exmple above. So I'm not going to repeate it. 

Instead I'm going to show you an example of how you might want to use it


Consider the output of `cellToBoundary()`

```{r}
h3r::cellToBoundary(cell = c("8cbe63562a54bff","8cbe635631103ff"))
```

This gives a list of 2 elements, and each element contains lat/lng coordinates.

You can use the C++ `h3r::cellToBoundary()` in your own workflow, and using other R packages that allow you to link to the source code.

In this example I'm suing `{geometries}` and `{sfheaders}` to convert to a valid `{sf}` object.

The steps inside this function are:

1. `h3r::cellToBoundary` - get the boundaries of each cell
2. `geometries::collapse_list()` - makes a single list, with three vectors
  - id
  - lat
  - lng
3. `sfheaders::sf_polygon()` - convert the result into an `sf` object


```{r}

library(Rcpp)
library(sf) ## for the `sf.print` method

cppFunction(
  
  depends = c("h3r", "geometries", "sfheaders")   # you need `sfheaders` installed
  , includes = c(
    '#include "geometries/utils/lists/collapse.hpp"'
    , '#include "sfheaders/sf/sf.hpp"'
    , '#include "h3rapi.h"'
  )
  
  , code = '
  
    SEXP sfBoundary(Rcpp::StringVector cells) {
      
      R_xlen_t n = Rf_xlength(cells);  
      
      // convert to latLng boundaries
      Rcpp::List boundaries = h3r::cellToBoundary(cells);
      
      // need to account for any pentagons 
      Rcpp::IntegerVector n_pentagons = h3r::isPentagon(cells);
      R_xlen_t n_pentagon = n_pentagons[0];
      R_xlen_t row_count = (n_pentagon * 5) + ((n - n_pentagon) * 6);
     
      // _collapse_ the boundaries to a list of three vectors
      // col0: id
      // col1: lat
      // col2: lng
      Rcpp::List geometries = geometries::utils::collapse_list(boundaries, row_count);
      
      // the `sfheaders` api expects a data.frame or a matrix
      Rcpp::DataFrame df = Rcpp::as< Rcpp::DataFrame >(geometries);

      Rcpp::IntegerVector idCol = {0};
      Rcpp::IntegerVector geometryCol = {1, 2};
      
      return sfheaders::api::sf_polygon(df, geometryCol, idCol, R_NilValue, "XY", false, true);

    }
  '
)


sfBoundary(cell = c("8cbe63562a54bff","8cbe635631103ff")) 

```





