---
title: "Core H3 Themes"
output: html_document
author: "Ray Shao, Sahinya Akila"
date: "2023-07-20"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## Core H3 Themes

There's lots of really [good content and documentation]() on the fundamentals of H3. This post touches on some of the core concepts, and shows you how you can use them in R through `h3r`

```{r}
library(h3r)
library(mapdeck)
library(data.table)
library(sfheaders)
```


> H3 is a geospatial indexing system that partitions the world into hexagonal cells
> 
> The H3 Core Library implements the H3 grid system. It includes functions for converting from latitude and longitude coordinates to the containing H3 cell, finding the center of H3 cells, finding the boundary geometry of H3 cells, finding neighbors of H3 cells, and more.

- [H3 v4. Introduction](https://h3geo.org/docs/)

## Cells

Resolution 0 cells are the [largest](https://h3geo.org/docs/core-library/restable#average-area-in-km2) (i.e, they cover the largest area) and resolution 15 are the smallest. The whole globe is covered with cells at each resolution. 

We're using the term 'cells' because each resolution is made up of _some_ number of hexagons, and exactly 12 pentagons.

If we plot all the Resolution 0 cells on a 'flat' map we get distortions as the cells get further from the equator. This is purely a result of plotting using the Mercator projection. However, all hexagons at each resolution are roughly the same size.

```{r resolution0}
mapdeck::set_token(secret::get_secret("MAPBOX"))

cells <- h3r::getRes0Cells()

mapdeck(libraries = "h3", style = mapdeck_style("dark"))  %>%
  add_h3(
    data = data.frame(x = cells)
    , hexagon = "x"
    , fill_colour = "x"
    , fill_opacity = 0.7
    , tooltip = "x"
    , stroke_colour = "#000000"
    , stroke_width = 50000
  )
```

#### Plotting (interlude)

We're using the `mapdeck` library for the maps because it uses [deck.gl](https://deck.gl/) as the underlying plotting engine. And given it is also developed by Uber, it can natively plot the H3 Cells, so there's no need to convert to lat/lon coordinates.

However, note that you have to tell `mapdeck()` you want to plot the cells by adding the `libraries = "h3"` parameter:

```{r eval=FALSE}

mapdeck(libraries = "h3") %>%
  add_h3()

```


### Pentagons

The reason for using pentagons is summarised by [Uber](https://h3geo.org/docs/core-library/overview/)

> The H3 grid is constructed on the icosahedron by recursively creating increasingly higher precision hexagon grids until the desired resolution is achieved. Note that it is impossible to tile the sphere/icosahedron completely with hexagons; each resolution of an icosahedral hexagon grid must contain exactly 12 pentagons at every resolution, with one pentagon centered on each of the icosahedron vertices. 

The nice feature about the pentagons is that they are all centered in the ocean.


```{r}
mapdeck::set_token(secret::get_secret("MAPBOX"))

cells <- h3r::getRes0Cells()
is_pentagon = h3r::isPentagon(cells)

mapdeck(libraries = "h3", style = mapdeck_style("dark"))  %>%
  add_h3(
    data = data.frame(x = cells, is_pentagon = is_pentagon)
    , hexagon = "x"
    , fill_colour = "is_pentagon"
    , fill_opacity = 0.5
    , tooltip = "x"
    , stroke_colour = "#000000"
    , stroke_width = 50000
  )
```

\

### Icosahedron faces

Those 122 resolution 0 cells live in 20 faces. Let's see which faces are they in.

If you hover over the cells you can notice that some intersect multiple faces. 

```{r}

cells <- h3r::getRes0Cells()
faces <- h3r::getIcosahedronFaces(cells)

## collapse the faces into a string of face IDs for each cell
faces <- sapply(faces, \(x) unlist(paste0(x, collapse = ","))) |> unname()

df_faces <- data.frame(
  cell = cells
  , face = faces
  )

mapdeck(libraries = "h3", style = mapdeck_style("dark"))  %>%
    add_h3(
        data = df_faces
        , hexagon = "cell"
        , fill_colour = "face"
        , fill_opacity = 0.7
        , tooltip = "face"
        , stroke_colour = "#000000"
        , stroke_width = 50000
    )
```

\

### Changing Resolution

If we plot all the Resolution 0 cells at the center of each icosahedron face, plus the children of each of these (by going up to resolution 1) we see there are (approximately) 7 children per cell.

```{r}
mapdeck::set_token(secret::get_secret("MAPBOX"))

cells <- c(
  '8021fffffffffff', '8005fffffffffff', '800ffffffffffff', '8035fffffffffff',
  '803ffffffffffff', '8065fffffffffff', '8033fffffffffff', '8049fffffffffff', '8081fffffffffff', '8097fffffffffff', '8073fffffffffff', '805dfffffffffff',
  '808ffffffffffff', '80c1fffffffffff', '80abfffffffffff', '80bffffffffffff',
  '80b5fffffffffff', '80d3fffffffffff', '80effffffffffff', '80e5fffffffffff'
  )

children_cells <- unlist(h3r::cellToChildren(cells, 1L))

mapdeck(libraries = "h3", style = mapdeck_style("dark"))  %>%
  add_h3(
    data = data.frame(x = c(cells, children_cells))
    , hexagon = "x"
    , fill_colour = "x"
    , fill_opacity = 0.5
    , tooltip = "x"
    , stroke_colour = "#000000"
    , stroke_width = 50000
  )
```

\

And if we continue to plot finer and finer resolution cells (all of which are children of the parent cell) we can start to see the rotation effect when changing resolution.

As you go from an odd-to-even resolution, the hexagons are rotated by ~19.1 degrees. Then as you go from even-to-odd, they are rotated back again.

So all the even-resolution cells are at the same orientation as each other, and similarly all the odd-resolution cells are at the same orientations as each other.

```{r}
mapdeck::set_token(secret::get_secret("MAPBOX"))

cells <- c('8081fffffffffff')

## Children and resolutions 1:4
children_cells <- unlist(h3r::cellToChildren(cells, 1L))
children_cells_res2 <- unlist(h3r::cellToChildren(children_cells[1], 2L))
children_cells_res3 <- unlist(h3r::cellToChildren(children_cells_res2[1], 3L))
children_cells_res4 <- unlist(h3r::cellToChildren(children_cells_res3[1], 4L))

mapdeck(
  libraries = "h3"
  , location = rev(as.numeric(cellToLatLng(cells)))
  , zoom = 3
  , style = mapdeck_style("dark")
  )  %>%
  add_h3(
    data = data.frame(
      x = c(cells, children_cells, children_cells_res2, children_cells_res3, children_cells_res4)
      )
    , hexagon = "x"
    , fill_colour = "x"
    , fill_opacity = 0.4
    , tooltip = "x"
    , stroke_colour = "#000000"
    , stroke_width = 5000
    , update_view = TRUE
  )
```

\

## Cells and Coordinates

For this section we'll be using the `stations` data from the h3r package

```{r}
( dt_stations <- as.data.table(h3r::stations) )
```

You can convert these coordinates to their containing cells...


```{r}
dt_stations[, cell := h3r::latLngToCell(lat = dt_stations$lat, lng = dt_stations$lon, resolution = 8) ][]
```

... but you can't get the coordinates back again. You can only get the center coordinates for each cell


```{r}
coords <- h3r::cellToLatLng(dt_stations$cell)
head(coords)

# Notice how they are different to head(stations)
```

### Boundaries

If you want the cell as a spatial polygon you can get their boundaries as coordinates, and convert to an `sf` object, as in this exmaple

```{r}

boundaries <- h3r::cellToBoundary(dt_stations$cell)

## The `cellToBoundary` function returns a named list, where each element
## is given the cell name. However, some stations are in the same cell
## dt_stations[, .N, by = cell ][ N != 1 ]
## so Replace the cell name with the stop name
names(boundaries) <- dt_stations$stop_name

boundaries <- data.table::rbindlist(boundaries, idcol = TRUE)

sf_boundaries <- sfheaders::sf_polygon(boundaries, x = "lng", y = "lat", polygon_id = ".id")
```

Notice in the map how some stations share the same cell (at resolution 8)

```{r}

mapdeck(
  location = c(144.967590, -37.810935)
  , zoom = 11
  ) %>% 
  add_scatterplot(
    data = dt_stations
    , update_view = FALSE
    , radius = 50
  ) %>%
  add_polygon(
    sf_boundaries
    , fill_colour = '#9f19ff88'
    , stroke_width = 25
    , update_view = FALSE
  )

```

\
\

## References

1. H3: Tiling the Earth with Hexagons [Youtube - Uber Engineering](https://www.youtube.com/watch?v=ay2uwtRO3QE&t=1487s)
2. Tables of Cell Statistics Across Resolutions [h3Geo.org](https://h3geo.org/docs/core-library/restable)

